---
title: "stats_final"
format: html
---

## Research Questions
1. Has the air quality (Nitrogen Dioxide levels) changed between Winter and Summer seasons?
2. Are there differences in Nitrogen Dioxide levels between selected neighborhoods in NYC?


```{r echo=FALSE}
#| message: false
#| warning: false
# Load libraries
library(tidyverse)
library(janitor)

# Load data
air_quality <- read_csv("data/Air_Quality.csv") %>%
  clean_names() # fixes column names to lowercase + underscores

# Filter NO2 and create Season column
no2_data <- air_quality %>%
  filter(name == "Nitrogen dioxide (NO2)") %>%
  mutate(
    season = case_when(
      str_detect(time_period, "Winter") ~ "Winter",
      str_detect(time_period, "Summer") ~ "Summer",
      TRUE ~ "Other"
    )
  ) %>%
  filter(season != "Other")

```


```{r echo=FALSE}
#| message: false
#| warning: false

ggplot(no2_data, aes(x = data_value)) +
  geom_histogram(binwidth = 2, fill = "skyblue", color = "black", na.rm = TRUE) +
  labs(title = "Distribution of NO₂ Levels", x = "NO₂ Level (ppb)", y = "Frequency") +
  theme_minimal()


```

```{r echo=FALSE}
#| message: false
#| warning: false
# Boxplot of NO2 Levels by Season
ggplot(no2_data, aes(x = season, y = data_value, fill = season)) +
  geom_boxplot() +
  labs(
    title = "NO₂ Levels by Season",
    x = "Season",
    y = "NO₂ Level (ppb)"
  ) +
  theme_minimal()


```

```{r echo=FALSE}
#| message: false
#| warning: false
# Summary statistics by Season
no2_data %>%
  group_by(season) %>%
  summarise(
    mean_NO2 = mean(data_value, na.rm = TRUE),
    sd_NO2 = sd(data_value, na.rm = TRUE),
    min_NO2 = min(data_value, na.rm = TRUE),
    max_NO2 = max(data_value, na.rm = TRUE)
  )


# Two-sample t-test comparing Winter and Summer NO2 levels
t_test_result <- t.test(data_value ~ season, data = no2_data)

# Print the results
t_test_result


```



## Conclusion

In this project, I analyzed air quality data from New York City, specifically focusing on Nitrogen Dioxide (NO₂) levels. The two research questions I explored were: (1) Has air quality changed between Winter and Summer seasons? and (2) Are there differences in NO₂ levels between selected neighborhoods? To address these questions, I conducted a descriptive analysis by creating a histogram of NO₂ levels and a boxplot comparing NO₂ levels across seasons. I also calculated summary statistics (mean, standard deviation, minimum, and maximum) for Winter and Summer seasons. To test for differences, I performed a two-sample t-test comparing NO₂ levels between Winter and Summer.

The results of the descriptive analysis suggested that there may be differences in air quality between seasons. The histogram showed the distribution of NO₂ levels, and the boxplot indicated that one season tended to have higher levels. The two-sample t-test yielded a p-value of 2.2e-16, indicating that there was a statistically significant difference between Winter and Summer NO₂ levels. Overall, the analysis suggests that seasonality has an impact on air quality in New York City, with Winter typically showing higher NO₂ concentrations.
